### 2017-03-08 Paper/Tables/Payment.tex

% Table generated by Excel2LaTeX from sheet 'payment_stats'
\begin{tabular}{rrrr}
\toprule
\multicolumn{1}{l}{Tipo Pago} & \multicolumn{1}{l}{N. Obs} & \multicolumn{1}{l}{Media} & \multicolumn{1}{l}{Desv. Est.} \\
\midrule
\multicolumn{4}{c}{Actor} \\
\midrule
\multicolumn{1}{l}{Pago para iniciar el juicio} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{6333.33} & \multicolumn{1}{c}{3511.88} \\
\multicolumn{1}{l}{Porcentaje(\%) de lo que obtengo en el juicio} & \multicolumn{1}{c}{95} & \multicolumn{1}{c}{29.69} & \multicolumn{1}{c}{12.27} \\
\multicolumn{1}{l}{Pago mensual} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{50500.00} & \multicolumn{1}{c}{70003.57} \\
\multicolumn{1}{l}{Un solo pago por el juicio} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{35.00} & \multicolumn{1}{c}{} \\
\multicolumn{1}{l}{Un solo pago SOLO si gano el juicio} & \multicolumn{1}{c}{13} & \multicolumn{1}{c}{8236.08} & \multicolumn{1}{c}{14419.82} \\
\multicolumn{1}{l}{Otra forma de pago} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
\multicolumn{1}{l}{Ningun pago} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
\midrule
\multicolumn{4}{c}{Representante Actor} \\
\midrule
\multicolumn{1}{l}{Pago para iniciar el juicio} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{3500.00} & \multicolumn{1}{c}{3000.00} \\
\multicolumn{1}{l}{Porcentaje(\%) de lo obtenido en el juicio} & \multicolumn{1}{c}{291} & \multicolumn{1}{c}{29.98} & \multicolumn{1}{c}{7.64} \\
\multicolumn{1}{l}{Pago mensual} & \multicolumn{1}{c}{7} & \multicolumn{1}{c}{8785.86} & \multicolumn{1}{c}{7381.41} \\
\multicolumn{1}{l}{Un solo pago por el juicio} & \multicolumn{1}{c}{8} & \multicolumn{1}{c}{11662.50} & \multicolumn{1}{c}{20019.91} \\
\multicolumn{1}{l}{Un solo pago SOLO si gana mi representado} & \multicolumn{1}{c}{25} & \multicolumn{1}{c}{3253.48} & \multicolumn{1}{c}{7136.72} \\
\multicolumn{1}{l}{Otra forma de pago} & \multicolumn{1}{c}{7} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
\multicolumn{1}{l}{Ningún pago} & \multicolumn{1}{c}{39} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} \\
\midrule
\multicolumn{4}{c}{Representante Demandado} \\
\midrule
\multicolumn{1}{l}{Pago por juicio} & \multicolumn{1}{c}{66} & \multicolumn{1}{c}{11717.61} & \multicolumn{1}{c}{12943.36} \\
\multicolumn{1}{l}{Pago en caso de ganar el juicio} & \multicolumn{1}{c}{17} & \multicolumn{1}{c}{10538.82} & \multicolumn{1}{c}{16118.45} \\
\multicolumn{1}{l}{Pago mensual (iguala)} & \multicolumn{1}{c}{209} & \multicolumn{1}{c}{13073.35} & \multicolumn{1}{c}{44400.94} \\
\multicolumn{1}{l}{Soy empleado de la empresa} & \multicolumn{1}{c}{58} & \multicolumn{1}{c}{10000.00} & \multicolumn{1}{c}{} \\
\multicolumn{1}{l}{Otro} & \multicolumn{1}{c}{90} & \multicolumn{1}{c}{15000.00} & \multicolumn{1}{c}{} \\
\bottomrule
\end{tabular}%