### 2017-03-08 Paper/working.tex

 \documentclass[11pt]{article}
\input{preamble.tex}

%\usepackage{refcheck}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DOCUMENT
\begin{document}

\title{Preliminary results for: Overconfidence and settlement: evidence from Mexican labor courts\thanks{We would like to thank Sebastian Garcia, Isaac Meza, Diana Roman, and Monica Zamudio (``la banda pesada'' and ``los mas aca'') for superhuman research assistance.  All errors are ours.}}
\author{Joyce Sadka \and Enrique Seira  \and Christopher Woodruff }
\date{This draft: \today \\[2 cm]}

%\vspace{.5in}

\maketitle

\section{Preliminary results}

\begin{landscape}

\subsection{Raw Expectations}

 \begin{table}[H]
 \caption{Belief updating winsorizing at 95 percentile}
 \label{Belief_updating}
 \begin{center}
 \scriptsize{\input{./Tables/exit_vs_initial_exp_95.tex}}
 \end{center}
  \footnotesize
 \textit{Notes:} 
 Dependent variable are expectations measured in the exit survey (for amount and delay respectively). Specification (2) includes basic controls.
 \textit{Do file: } \texttt{exit\_vs\_initial\_exp.do}
 \end{table}

 \begin{table}[H]
 \caption{Belief updating winsorizing at 90 percentile}
 \label{Belief_updating}
 \begin{center}
 \scriptsize{\input{./Tables/exit_vs_initial_exp_90.tex}}
 \end{center}
  \footnotesize
 \textit{Notes:} 
 Dependent variable are expectations measured in the exit survey (for amount and delay respectively). Specification (2) includes basic controls.
 \textit{Do file: } \texttt{exit\_vs\_initial\_exp.do}
 \end{table}

 \begin{table}[H]
 \caption{Belief updating (amounts in log)}
 \label{Belief_updating_log}
 \begin{center}
 \scriptsize{\input{./Tables/exit_vs_initial_exp_log.tex}}
 \end{center}
  \footnotesize
 \textit{Notes:} 
 Dependent variable are expectations measured in the exit survey (for amount (in log pesos) and delay respectively). Specification (2) includes basic controls.
 \textit{Do file: } \texttt{exit\_vs\_initial\_exp\_log.do}
 \end{table}

\subsection{Relative OC Expectations}

 \begin{table}[H]
 \caption{Belief updating (relative) winsorizing at 95 percentile}
 \label{Belief_updating}
 \begin{center}
 \scriptsize{\input{./Tables/exit_vs_initial_exp_95_rel.tex}}
 \end{center}
  \footnotesize
 \textit{Notes:} 
 Dependent variable are expectations measured in the exit survey (for amount and delay respectively). Specification (2) includes basic controls.
 \textit{Do file: } \texttt{exit\_vs\_initial\_exp\_rel.do}
 \end{table}

 \begin{table}[H]
 \caption{Belief updating (relative) winsorizing at 90 percentile}
 \label{Belief_updating}
 \begin{center}
 \scriptsize{\input{./Tables/exit_vs_initial_exp_90_rel.tex}}
 \end{center}
  \footnotesize
 \textit{Notes:} 
 Dependent variable are expectations measured in the exit survey (for amount and delay respectively). Specification (2) includes basic controls.
 \textit{Do file: } \texttt{exit\_vs\_initial\_exp\_rel.do}
 \end{table}

 \begin{table}[H]
 \caption{Belief updating (relative) (amounts in log)}
 \label{Belief_updating_log}
 \begin{center}
 \scriptsize{\input{./Tables/exit_vs_initial_exp_log_rel.tex}}
 \end{center}
  \footnotesize
 \textit{Notes:} 
 Dependent variable are expectations measured in the exit survey (for amount (in log pesos) and delay respectively). Specification (2) includes basic controls.
 \textit{Do file: } \texttt{exit\_vs\_initial\_exp\_log\_rel.do}
 \end{table}

\end{landscape}

\subsection{Private Lawyer interaction}
% INTERACTION WITH PRIVATE LAWYER
\begin{table}[H]
    \caption{Treatment Efects by Month}
    \label{Table_effects}
    \begin{center}
    \scriptsize{\input{./Tables/Minicortes_2_int.tex}}
    \end{center}
    \footnotesize
    \textit{Notes:} 
    Calc=Conc and Calc=Conc=0 shows the p-value for the respective tests.
    \textit{Do file: } \texttt{mini\_cortes\_2\_int.do}
\end{table}

\begin{table}[H]
    \caption{Triple interaction}
    \label{Table_effects}
    \begin{center}
    \scriptsize{\input{./Tables/Minicortes_3int.tex}}
    \end{center}
    \footnotesize
    \textit{Notes:} 
    Calc=Conc and Calc=Conc=0 shows the p-value for the respective tests.
    \textit{Do file: } \texttt{mini\_cortes\_3int.do}
\end{table}

\begin{table}[H]
  \centering
  \caption{Tab presence of employee and public lawyer}
\begin{tabular}{rccr}
\toprule
      & \multicolumn{2}{c}{Employee presence} &  \\
\midrule
\multicolumn{1}{c}{Public Lawyer} & 0     & 1     & \multicolumn{1}{c}{Total} \\
\multicolumn{1}{c}{0} & 666   & 137   & \multicolumn{1}{c}{803} \\
\multicolumn{1}{c}{1} & 51    & 38    & \multicolumn{1}{c}{89} \\
\multicolumn{1}{c}{Total} & 717   & 175   & \multicolumn{1}{c}{892} \\
\bottomrule
\end{tabular}%

\end{table}%

\end{document}