### 2017-06-28 ScaleUp/ITT_ATT.tex
replace png to tiff

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{lscape}
\usepackage{tikz}
\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{longtable}

\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}

\newtheorem{theorem}{Teorema}
\newtheorem{lem}[theorem]{Lemma}
\newtheorem{dfn}{Definición}
\newtheorem{cor}[theorem]{Corolario}
\newtheorem{obs}{Obs}
\newtheorem{rem}{Remark}
\newtheorem{prob}{Problema}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=.05pt] (char) {#1};}}

\newtheoremstyle{named}{}{}{\itshape}{}{\bfseries}{.}{.5em} {\thmnote{#3 }#1}
\theoremstyle{named}
\newtheorem*{namedtheorem}{}

\newcounter{exercisecounter}
\newenvironment{ex}{\begin{quote}%
    \refstepcounter{exercisecounter}%
  \textbf{Ejercicio \arabic{exercisecounter}}%
  \quad
}{%
\end{quote}%
}
\newcounter{ejemplocounter}
\newenvironment{ej}{\begin{quote}%
    \refstepcounter{ejemplocounter}%
  \textbf{Ejemplo \arabic{ejemplocounter}}%
  \quad
}{%
\end{quote}%
}

\renewcommand{\d}[1]{\ensuremath{\operatorname{d}\!{#1}}}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

\newcommand{\folder}{./Effect}

\begin{document}

\title{Treatment effects}

\author{Instituto Tecnológico Autónomo de México}
\date{\today}
\maketitle

\hrulefill

\section{Some Summary Statistics}

\vspace{7mm}

\subsection*{Administrative Stats}

Summary statistics of basic subject characteristics. All variables coming from casefile details are tested for balance between survey and non-survey populations.

\begin{center}
\scriptsize{\input{\folder/admin_stats.tex}}
\end{center}

\pagebreak

\begin{center}
\scriptsize{\input{\folder/TAB_survey.tex}}
\end{center}

\pagebreak

\begin{landscape}
\begin{center}
\scriptsize{\input{\folder/ITT_ATT_report.tex}}
\end{center}

\end{landscape}

\subsection*{Expectation Stats}

Summary statistics for expectations.

\begin{center}
\scriptsize{\input{\folder/expectation_stats.tex}}
\end{center}

\pagebreak

\subsection*{Update in beliefs}

The update in beliefs is measured in relative terms, this is the percentage deviation of the exit survey with respect to the entrance survey
\[\frac{exit-initial}{initial}\]

We provide two measures: Update in beliefs in probability and payment.\\

\begin{center}
\scriptsize{\input{\folder/update_beleif.tex}}
\end{center}

\begin{figure}[H]
\label{diff}
\begin{center}
\includegraphics[width=\textwidth]{./Figures/belief.pdf}
\end{center}
{\footnotesize \textit{The histograms are trimmed at the 75 percentile in the case for amount. Width of bins are \$30,000 pesos for the case of amount and 10\% for the case of probability.}}
\end{figure}

\begin{figure}[H]
\label{update}
\begin{center}
\includegraphics[width=\textwidth]{./Figures/update_belief.pdf}
\end{center}
{\footnotesize \textit{The histograms measures the update in relative terms. It displays the distribution from the percentile 5 to 95.}}
\end{figure}

\pagebreak

\begin{figure}[H]
\label{update}
\begin{center}
\includegraphics[width=\textwidth]{./Figures/updatebeleif_amount.pdf}
\end{center}
{\footnotesize \textit{Notes: Graphs on belief updating. They show the following statistic: $\theta=\left|\frac{P-exitsurvey}{P-initialsurvey}\right|$. Where $P$ is the prediction made by the calculator and $exit$ and $initial$ survey variables denote what each part answered on the survey for amount won. Note that when $\theta=1$ we have no update with respect to the prediction. And when $\theta<1$ they update in the direction of what calculator says. The extreme case $\theta=0$ implies perfect updating with respect to what the calculator says. }}
{\footnotesize \textit{Do file: } \texttt{Update\_beleifs.do}}
\end{figure}

\pagebreak

The following table displays a regression of conciliation against the statistics of updating
\[\theta=\left|\frac{P-exitsurvey}{P-initialsurvey}\right|\]

\begin{center}
\scriptsize{\input{\folder/con_vs_update.tex}}
\end{center}

Now we test difference in initial and exit survey in expectations. The following table shows the results.

\begin{center}
\scriptsize{\input{\folder/ttest.tex}}
\end{center}

\pagebreak

\section{ITT and ATT regressions}

\begin{landscape}
\subsection*{ITT and Second Stage ATT}

\begin{center}
\scriptsize{\input{\folder/ITT_ATT.tex}}
\end{center}

\end{landscape}

\pagebreak

\subsection*{First Stage for ATT regressions}

\begin{center}
\scriptsize{\input{\folder/FS_ATT.tex}}
\end{center}

\section{Take up regression (calculator and survey)}

\begin{center}
\scriptsize{\input{\folder/Take_up.tex}}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}