### 2017-07-19 Pilot_3/main.tex
Do files updated for pilot 3

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amscd}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{lscape}
\usepackage{tikz}
\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white
\usepackage{fullpage}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{longtable}

\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

\newcommand{\folder}{./Tables}
\graphicspath{{./Figuras/}}

\begin{document}

\title{Pilot 3}

\author{Instituto Tecnológico Autónomo de México}
\date{\today}
\maketitle

\hrulefill

\subsection*{Balance Table}

\begin{figure}[H]
    \caption{Cumulative number of cases by date/treatment}
    \label{}
    \begin{center}
    \begin{subfigure}{0.49\textwidth}
    \centering
        \includegraphics[width=\textwidth]{cum_num_cases.pdf}
     \end{subfigure}
     \begin{subfigure}{0.49\textwidth}
    \centering
        \includegraphics[width=\textwidth]{cum_num_cases_tr.pdf}
     \end{subfigure}
  \end{center}
        {\footnotesize \textit{Notes:Cumulative number of cases by date and treatment.}}
         {\footnotesize \textit{Do file: } \texttt{cum\_num\_cases.do}}
\end{figure}

\begin{table}[H]
\caption{Balance table}
\begin{center}
\scriptsize{\input{./Tables/Balance.tex}}
\end{center}
 \footnotesize
\textit{Notes: Balance table on some variables. Rightmost columns shows p-value for a multivariate tests of means between different arm treatments.} 
\textit{Do file: } \texttt{balance.do}
\end{table}

\begin{table}[H]
\caption{Balance table from 2017-06-15}
\begin{center}
\scriptsize{\input{./Tables/Balance_c.tex}}
\end{center}
 \footnotesize
\textit{Notes: Balance table on some variables. Rightmost columns shows p-value for a multivariate tests of means between different arm treatments.} 
\textit{Do file: } \texttt{balance.do}
\end{table}

\begin{table}[H]
\caption{Balance table with july data}
\begin{center}
\scriptsize{\input{./Tables/Balance_july.tex}}
\end{center}
 \footnotesize
\textit{Notes: Balance table on some variables. Rightmost columns shows p-value for a multivariate tests of means between different arm treatments.} 
\textit{Do file: } \texttt{balance.do}
\end{table}

\end{document}