### 2017-08-10 presentations/Tables/Experiment_description.tex
Presentations folder

% Table generated by Excel2LaTeX from sheet 'Experiment_description'
\begin{tabular}{cccc}
\toprule
      & Control & Calculator & Conciliator \\
\midrule
Baseline survey & \multicolumn{3}{c}{All treatment and control groups were required to take baseline survey} \\
Calculator prediction & -     & Subjects were assisted by project personnel to input variables from their case file into the calculator, and the resulting predictions were then explained. & Subjects did not have access to the calculator information \\
Conciliator mediation & -     & Subjects could choose to talk to a conciliator.    & Subjects were required to talk to the conciliator, whether their counterpart was present or not. \\
Exit survey & \multicolumn{3}{c}{All treatment and control groups were required to take exit survey} \\
\bottomrule
\end{tabular}%